Reasoning About a Machine with Local Capabilities

### 2018-05-11 toplas/paper.tex
Finished pass over proof sketch.

  The capability for the closure is a global enter capability that we call $c_{f4}$.
  The remainder of the proof amounts to showing that it is safe to return $c_{f4}$ to the adversary, i.e.
  $c_{f4} \in \stdvr(W')$ where $W'$ is $W$ with the regions added after executing \texttt{g1}.
  The adversary can use $c_{f4}$ whenever, so all we may assume about the configuration that $c_{f4}$ is invoked in is that the register-file $\reg$ and memory $\ms$ satisfies a world $W_1$ that is a private future world of $W'$, i.e.\ $\reg \in \stdrr(W_1)$ and $\memSat[]{\ms}{W_1}$.
  We have to show that the invocation respects $W_1$ which corresponds to showing $(\reg[\pcreg \mapsto \updatePcPerm{c_{f4}}],\ms) \in \observations(W_1)$.
  When $c_{f4}$ is invoked, we know exactly what instructions are executed up until the \texttt{scall}, so we can use the anti-reduction lemma (Lemma~\ref{lem:anti-red-obs}) to reason about this part.
  During the first bit of execution, it is checked that the stack pointer has permission read/write-local/execute and that the callback is global.
  Based on the stack pointers permission, we know that the capability is local which means that the region that governs it must be temporary.
  This allows us to construct $W_2 \futurestr W_1$ where the stack region has been revoked (notice that it is also important for the reasoning about $x$ that this is a private future world).
  In $W_2$, we replace the stack region with a temporary static region that governs our stack frame including the private registers we push to the stack and the stack recovery code as well as a standard region $\iota^{pwl}$-region for the remainder of the stack.
  As the callback is global, it is safe to invoke it in a private future world of $W_1$ (even though the code is unknown, we do not need to use the FTLR because we may assume that the arguments for the invocation of $c_{f4}$ are safe).
  Notice that had the callback been local, this would not be the case.
  We also need to reason about the return pointer we provide in the invocation.
  The return pointer is a local capability, so we may assume that it is invoked in some $W_3 \futurewk W_2$.
  This means that none of the temporary regions have been revoked, so the regions we replaced the old stack region with are still there in $W_3$.
  For the sake of presentation, we went through the above reasoning, but we actually us Lemma~\ref{lem:scall-works} to reason about \texttt{scall}.
  To reason about the next bit of execution we know, we once again use Lemma~\ref{lem:anti-red-obs} followed by \ref{lem:scall-works} to handle \texttt{scall}.
  As the private stack has changed, we need to replace the temporary static region that governs it, so it is revoked.
  This means that the \texttt{scall} takes place in a world $W_4 \futurestr W_3$.
  The reasoning about \texttt{scall} is the same as for the first callback invocation, so again we will get a world $W_5 \futurewk W_4$ in which we need to argue that the remainder of the execution is safe.
  At this point, we can use Lemma~\ref{lem:anti-red-obs} to get to the point where we jump to the return pointer.
  If we can argue that it is safe to jump to the return pointer we got initially from the adversary, then the proof is done.
  We have made no checks on the return pointer, so we have no idea whether it is local or global\footnote{We may assume that it is a capability that is executable when jumped to otherwise the execution fails which is considered safe.}.
  W.l.o.g.\ assume the return pointer is local.
  This means that we have to use it in a world $W_6$ that is a public future world of $W_1$, which means that we have to restore all the regions in the world that the adversary relies on.
  In particular, we will have to restore the stack region from $W_1$.
  We cannot just pick $W_1$ because Lemma~\ref{lem:anti-red-obs} requires $W_6 \futurestr W_5$.
  Before we proceed, let us consider where each of the worlds came from so far: As illustrated in \figurename~\ref{fig:worlds-in-pf-g1}, we chose $W_2$ and $W_4$ which means that we know exactly what changes they made.
  In particular, we did not mask any of the revoked or temporary regions in $W_1$ with a permanent region.
  This means that in $W_6$, we can reinstate every temporary and revoked region in $W_1$.
  In the private future worlds we construct, we only take transitions that relative to $W_1$ are public.
  The worlds we are given, $W_3$ and $W_5$ may have taken arbitrary public transitions, but this is not problem with respect to the public future world relation.
  In $W_5$ there will be regions to handle the stack.
  These regions obviously need to be revoked as the $W_1$ stack regions replace them.
  The non-stack regions in $W_5$ that are not present in $W_1$ are also added in $W_6$, which the public future world permits as it is extensional.
  All in all, we create a world in which the invariants that the adversary relies on has been restored which makes it safe to invoke the return pointer.

@@

For example, an untrusted advertisement scenario with initialization code
that registers a redraw callback. The redraw callback gets temporary
read-write access to a framebuffer.

>>>>

  The capability for the closure is a global enter capability that we call $c_{f4}$.
  The remainder of the proof amounts to showing that it is safe to return $c_{f4}$ to the adversary, i.e.
  $c_{f4} \in \stdvr(W')$ where $W'$ is $W$ with the regions added after executing \texttt{g1}.
  The adversary can use $c_{f4}$ whenever, so all we may assume about the configuration that $c_{f4}$ is invoked in is that the register-file $\reg$ and memory $\ms$ satisfies a world $W_1$ that is a private future world of $W'$, i.e.\ $\reg \in \stdrr(W_1)$ and $\memSat[]{\ms}{W_1}$.
  We have to show that the invocation respects $W_1$ which corresponds to showing $(\reg[\pcreg \mapsto \updatePcPerm{c_{f4}}],\ms) \in \observations(W_1)$.
  When $c_{f4}$ is invoked, we know exactly what instructions are executed up until the \texttt{scall}, namely it is ensured that the callback is global and the stack pointer has read/write-local/execute permission, and $x$ is set to 0.
  Because we know part of the execution, we can apply the anti-reduction lemma (Lemma~\ref{lem:anti-red-obs}).

  The next part in the execution is an \texttt{scall}, so according to \sectionname~\ref{sec:reasoning}, we should apply the \texttt{scall} lemma (Lemma~\ref{lem:scall-works}).
  For the sake of presentation, we here sketch some of the things the \texttt{scall} lemma actually takes care of.
  Based on the stack pointer's permission, we know that the capability is local (because it is a system wide assumption that there are no global read/write-local/execute capabilities \textit{c.f.}\ \sectionname~\ref{sec:stack-and-return-pointer}) which means that the region that governs it must be temporary.
  This allows us to construct a world $W_2 \futurestr W_1$ where the stack region of $W_1$ has been revoked\footnote{As a side note, it is also important for the reasoning about $x$ that this is a private future world.}.
  In $W_2$, we replace the stack region of $W_1$ with a temporary static region that governs our stack frame, that is the private registers we push to the stack and the stack recovery code, as well as a standard region $\iota^{pwl}$-region for the part of the stack, we will provide to the callback.
  The static region ensures that our local stack frame remains unchanged during the execution of the callback, and the standard region makes sure that the callback only puts safe values on the stack.
  The callback is global, so it is safe to invoke it in a private future world of $W_1$ (even though the code is unknown, we do not need to use the FTLR because we may assume that the arguments for the invocation of $c_{f4}$ are safe).
  Notice that had the callback been local, this would not be the case, but it would also not be safe to invoke as it might be a stack pointer as discussed in \sectionname~\ref{sec:stack-and-return-pointer}.
  This part corresponds to arguing about Hyp-Callee in the \texttt{scall} lemma.

  We need to argue that it is safe to give the return pointer that \texttt{scall} constructs to the adversary which correspond to reason about the remainder of the execution of \texttt{f4} (corresponding to the Hyp-Cont case of the \texttt{scall} lemma).
  The return pointer is a local capability, so we may assume that it is invoked in some configuration that satisfies $W_3 \futurewk W_2$.
  This means that none of the temporary regions have been revoked, so the regions we replaced the old stack region with are still present in $W_3$.
  We know exactly how the execution proceeds upon return (the recovery code is executed, and $x$ is set to 0), so for the part of the execution up until the \texttt{scall} we once again use the anti-reduction lemma.

  For the \texttt{scall} itself we of course apply the \texttt{scall} lemma.
  As the private stack has changed, we need to replace the two regions that govern the stack.
  This means that the second invocation of the callback takes place in a world $W_4 \futurestr W_3$.
  The reasoning about \texttt{scall} is the same as for the first callback invocation.
  The callback capability is still safe because it is global which basically covers the Hyp-Callee case.
  For the Hyp-Cont case, we get a world $W_5 \futurewk W_4$ in which we need to argue that the remainder of the execution is safe.
  At this point, we can use the anti-reduction lemma a final time to get to the point where we jump to the return pointer.

  If we can argue that it is safe to jump to the return pointer that we got initially from the adversary, then the proof is done.
  We have made no checks on the return pointer, so we have no idea whether it is local or global\footnote{We may assume that it is a capability that is executable when jumped to otherwise the execution fails which is considered safe.}.
  W.l.o.g.\ assume the return pointer is local.
  This means that it is safe in public future worlds of $W_1$ (remember that it came from $\reg$ and $\reg \in \stdrr(W_1)$).
  This means that we need to construct a world $W_6$ with all the temporary and revoked regions of $W_1$ (this corresponds to restoring the invariants the adversary relies on for its safe execution).
  Further for all these regions, $W_6$ must use the same region names as $W_1$ for $W_6$ to be a public future world.
  As we used the anti-reduction lemma to get to this point, $W_6$ must also be constructed to be a private future world of $W_5$.
  Before we construct $W_6$, let us consider where each of the worlds we have seen so far came from: As illustrated in \figurename~\ref{fig:worlds-in-pf-g1}, we constructed $W_2$ and $W_4$ which are the only private future worlds we have considered so far.
  This means that we know exactly what changes they made.
  In particular, we did not mask any of the temporary or revoked regions in $W_1$ with a permanent region.
  This means that in $W_6$, we can reinstate every temporary or revoked region of $W_1$.
  In the private future worlds we constructed, we only took transitions that, relative to $W_1$, are public.
  The worlds we were given, $W_3$ and $W_5$ may have taken arbitrary public transitions, but this is no problem with respect to the public future world relation.
  In $W_5$ there were regions to handle the stack.
  These regions obviously need to be revoked as the $W_1$ stack regions replace them.
  The non-stack regions in $W_5$ that are not present in $W_1$ are also added in $W_6$, which the public future world relation permits as it is extensional.
  All in all, it is possible to construct $W_6$, so it is both a public future world of $W_1$ and a private world of $W_5$ which means that it is indeed safe to return to the adversary.

@@

For example, an untrusted advertisement scenario with initialization code
that registers a redraw callback. The redraw callback gets temporary
read/write access to a framebuffer.
